\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm, graphics,enumerate,fancyhdr}

\pagestyle{fancy}


\setlength{\parindent}{0pt}
\setlength{\textwidth}{6.5in}
\setlength{\oddsidemargin}{0in}
\addtolength{\textheight}{1in}

\renewcommand\theenumi{\alph{enumi}}
\renewcommand\labelenumi{(\theenumi)}


\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\CC}{\mathbb{C}}
\renewcommand\vec{\mathbf}

\theoremstyle{definition}
\newtheorem{problem}{}
\newtheorem{solution}{}

\begin{document}
\fancyhf{} % sets both header and footer to nothing
\renewcommand{\headrulewidth}{0pt}
\fancyfoot[L]{Copyright 2024 The University of British Columbia}

%\pagestyle{empty}

\begin{center} 
{\bf{Math 321 Assignment 0}} \\ 
{\bf{Due Friday, January 12 at 9 am}} 
\end{center}  
\bigskip
\hrule
\begin{center}
{\underline{Instructions}} 
\end{center} 
\begin{enumerate}[(i)] 
\item Homework should be submitted using Canvas. Include your name and SID.
%
\item Solutions should be well-crafted, legible and written in complete English sentences. You will be graded both on accuracy as well as the quality of exposition.
%
\item Theorems stated in the text or proved in lecture do not need to be reproved. Any other statement should be justified rigorously.
% 
\end{enumerate} 
\hrule
\bigskip

The purpose of this homework is to make sure you can use Canvas to submit homework. This homework is not worth any credit, but you should do it anyway---Homework 1 \emph{is} worth credit, so you should be sure you know how to submit using Canvas before Homework 1 is due. 

\begin{problem}
What is your name?
\end{problem}

\begin{problem}
What is your quest?
\end{problem}

\begin{problem}
What is your favourite colour?
\end{problem}

\begin{problem}
[Bonus] What is the capital of Assyria?
\end{problem}
\end{document}

\end{document}
